Coupling finite and boundary element methods to solve the Poisson-Boltzmann equation for electrostatics in molecular solvation

### 2023-04-27 conclusion.tex
Updated Results with tol 1e-4

This paper presents the first implementation of a FEM-BEM coupling approach to solve the Poisson-Boltzmann equation for molecular electrostatics. This brings the best of both worlds: the accuracy and efficiency of BEM to exactly enforce the boundary conditions at infinity, and the flexibility of FEM to account for space variations of the material properties and nonlinearities. After presenting verification results for a sphere and arginine with a constant permittivity inside the solute, we showcased our implementation with an advanced modeling technique that considers a Gaussian-varying permittivities in a confined region, comparing with the widely-used APBS software. The final scaling results for larger molecules start from protein G B1 (955 atoms) and go up to immunoglobulin G (20 148 atoms), proving the applicability of this approach for realistic problems. Even though our implementation was able to reach medium-to-large systems, we recongnize the need of further research towards opmtimizing the coupling technique, acceleration algorithms, and parallel execution, specially as we look towards much larger solutes, like viruses. We hope this proof-of-concept work will serve as motivation for future model development that considers space varying permittivities and Debye lengths, and mixed linear-nonlinear techniques, specially for highly-charged systems, like nucleic acids.

>>>>

This paper presents the first implementation of a FEM-BEM coupling approach to solve the Poisson-Boltzmann equation for molecular electrostatics. This brings the best of both worlds: the accuracy and efficiency of BEM to exactly enforce the boundary conditions at infinity, and the flexibility of FEM to account for space variations of the material properties and nonlinearities. After presenting verification results for a sphere and arginine with a constant permittivity inside the solute, we showcased our implementation with an advanced modeling technique that considers Gaussian-varying permittivities in a confined region, compared with the widely-used APBS software. The final scaling results for larger molecules start from protein G B1 (955 atoms) and go up to immunoglobulin G (20 148 atoms), proving the applicability of this approach for realistic problems. Even though our implementation was able to reach medium-to-large systems, we recongnize the need for further research towards optimizing the coupling technique, acceleration algorithms, and parallel execution, especially as we look towards much larger solutes, like viruses. We hope this proof-of-concept work will serve as motivation for future model development that considers space-varying permittivities and Debye lengths, and mixed linear-nonlinear techniques, especially for highly-charged systems, like nucleic acids.

### 2023-04-27 intro.tex
Updated Results with tol 1e-4

In biologically relevant settings, the structure and function of biomolecules is largely determined by the surrounding water, which usually contains salt. 
To describe these systems accurately, we need to account for the solvent correctly, which has given rise to a wide range of models.\cite{onufriev2018water}
Highly detailed models consider every water molecule and salt ion explicitly, however, there are approximated models that use continuum theory to represent this ionic solution, known as implicit-solvent models.\cite{RouxSimonson1999,DecherchiETal2015}
In the case of electrostatics, the implicit-solvent model is mathematically characterized by the Poisson-Boltzmann equation (PBE)\cite{Baker2004,Bardhan2012}, which is widely used to compute solvation free energies and mean-field potentials.

The implicit-solvent model for electrostatics describes the dissolved molecule as a infinite medium with a low-dielectric solute-shaped cavity, which contains a charge distribution from the partial charges --- usually a sum of Dirac deltas at the atom's locations.
The outer solvent region is represented with a high dielectric, and considers the presence of salt.
These two regions are interfaced by the molecular surface, which can be defined in various ways \cite{HarrisBoschitcshFenley2013}, where the continuity of the electrostatic potential and electric displacement are enforced.

The PBE has been solved numerically with finite difference\cite{BakerETal2001,GilsonETal1985,JurrusETal2018,LiETal2019}, finite element\cite{HolstETal2012,BondEtal2010}, boundary element\cite{boschitsch2002fast,LuETal2006,AltmanBardhanWhiteTidor09,bajaj2011efficient,GengKrasny2013,CooperBardhanBarba2014}, and analytical\cite{YapHeadgordon2010,FelbergETal2017} methods.
In particular, the boundary element method (BEM) has proven to be very efficient for high accuracy calculations \cite{GengKrasny2013,CooperBardhanBarba2014}, mainly due to the precise description of the molecular surface and point charges. 
However, BEM is limited to constant material properties in each region, and the linear version of the PBE. 
Even though these limitations are acceptable in a wide range of applications, there are cases when BEM falls short, for example, if a variable permittivity is required inside the solute \cite{grant2001smooth,li2013dielectric}, or the solute is highly charged such that the linear approximation breaks\cite{FogolariETal1999}.

The present article describes a methodology to overcome some of those limitations, by coupling finite and boundary elements methods.
This approach brings the best of both worlds: the flexibility of FEM and efficiency of BEM, all in an accurate description of the solute molecule.
FEM-BEM coupling is a popular technique in the context of mixed linear-nonlinear models,\cite{carstensen1995coupling,aurada2013classical} fracture mechanics,\cite{aour2007coupled} fluid-structure interaction,\cite{estorff1991fem} acoustics,\cite{hiptmair2006stabilized} and electromagnetics.\cite{matsuoka1988calculation,hiptmair2008stabilized,bruckner20123d}
On the other hand, the PBE has been solved with hybrid numerical methods in the past, for example, by coupling finite difference with boundary elements\cite{boschitsch2004hybrid} or finite elements,\cite{xie2016new,ying2018hybrid} which were mainly used to solve the nonlinear PBE in a specific region, and to implement modifications to the PBE model ({\it ie.} the size-modified PBE).
To the best of our knowledge, this is the first time finite and boundary elements are combined in this application.

Here we prototype this principle with the simplest implementation possible: a Johnson-N\'ed\'elec\cite{johnson1980coupling} coupling, mass-matrix and diagonal preconditioning, and no distributed memory execution. 
This limits the size of problems we can access currently, however, it sets the basis for future developments that use more elaborate formulations and algorithms that are readily available in the software libraries used in this work, in particular, Bempp-cl and FEniCSx.
These libraries are designed for ease-of-use through a simple Python interface, making them ideal tools to explore and improve computational efficiency, towards tackling large-scale problems, such as a full viral capsid.\cite{MartinezETal2019,wang2021high}
We hope this work will inspire research along these lines.

>>>>

In biologically relevant settings, the structure and function of biomolecules are largely determined by the surrounding water, which usually contains salt. 
To describe these systems accurately, we need to account for the solvent correctly, which has given rise to a wide range of models.\cite{onufriev2018water}
Highly detailed models consider every water molecule and salt ion explicitly, however, there are approximated models that use continuum theory to represent this ionic solution, known as implicit-solvent models.\cite{RouxSimonson1999,DecherchiETal2015}
In the case of electrostatics, the implicit-solvent model is mathematically characterized by the Poisson-Boltzmann equation (PBE)\cite{Baker2004,Bardhan2012}, which is widely used to compute solvation free energies and mean-field potentials.

The implicit-solvent model for electrostatics describes the dissolved molecule as an infinite medium with a low-dielectric solute-shaped cavity, which contains a charge distribution from the partial charges --- usually a sum of Dirac deltas at the atom's locations.
The outer solvent region is represented with a high dielectric, and considers the presence of salt.
These two regions are interfaced by the molecular surface, which can be defined in various ways \cite{HarrisBoschitcshFenley2013}, where the continuity of the electrostatic potential and electric displacement are enforced.

The PBE has been solved numerically with finite difference\cite{BakerETal2001,GilsonETal1985,JurrusETal2018,LiETal2019}, finite element\cite{HolstETal2012,BondEtal2010}, boundary element\cite{boschitsch2002fast,LuETal2006,AltmanBardhanWhiteTidor09,bajaj2011efficient,GengKrasny2013,CooperBardhanBarba2014}, and analytical\cite{YapHeadgordon2010,FelbergETal2017} methods.
In particular, the boundary element method (BEM) has proven to be very efficient for high-accuracy calculations \cite{GengKrasny2013,CooperBardhanBarba2014}, mainly due to the precise description of the molecular surface and point charges. 
However, BEM is limited to constant material properties in each region and the linear version of the PBE. 
Even though these limitations are acceptable in a wide range of applications, there are cases when BEM falls short, for example, if a variable permittivity is required inside the solute \cite{grant2001smooth,li2013dielectric}, or the solute is highly charged such that the linear approximation breaks\cite{FogolariETal1999}.

The present article describes a methodology to overcome some of those limitations, by coupling finite and boundary elements methods.
This approach brings the best of both worlds: the flexibility of FEM and the efficiency of BEM, all in an accurate description of the solute molecule.
FEM-BEM coupling is a popular technique in the context of mixed linear-nonlinear models,\cite{carstensen1995coupling,aurada2013classical} fracture mechanics,\cite{aour2007coupled} fluid-structure interaction,\cite{estorff1991fem} acoustics,\cite{hiptmair2006stabilized} and electromagnetics.\cite{matsuoka1988calculation,hiptmair2008stabilized,bruckner20123d}
On the other hand, the PBE has been solved with hybrid numerical methods in the past, for example, by coupling finite difference with boundary elements\cite{boschitsch2004hybrid} or finite elements,\cite{xie2016new,ying2018hybrid} which were mainly used to solve the nonlinear PBE in a specific region, and to implement modifications to the PBE model ({\it ie.} the size-modified PBE).
To the best of our knowledge, this is the first time finite and boundary elements are combined in this application.

Here we prototype this principle with the simplest implementation possible: a Johnson-N\'ed\'elec\cite{johnson1980coupling} coupling, mass-matrix and diagonal preconditioning, and no distributed memory execution. 
This limits the size of problems we can access currently, however, it sets the basis for future developments that use more elaborate formulations and algorithms that are readily available in the software libraries used in this work, in particular, Bempp-cl and FEniCSx.
These libraries are designed for ease of use through a simple Python interface, making them ideal tools to explore and improve computational efficiency, towards tackling large-scale problems, such as a full viral capsid.\cite{MartinezETal2019,wang2021high}
We hope this work will inspire research along these lines.

### 2023-04-27 main.tex
Updated Results with tol 1e-4

\begin{abstract}
The Poisson-Boltzmann equation is widely used to model electrostatics in molecular systems. Available software packages solve it using finite difference, finite element, and boundary element methods, where the latter is attractive due to the accurate representation of the molecular surface and partial charges, and exact enforcement of the boundary conditions at infinity. However, the boundary element method is limited to linear equations and piece-wise constant variations of the material properties. In this work, we present a scheme that couples finite and boundary elements for the Poisson-Boltzmann equation, where the finite element method is applied in a confined {\it solute} region, and the boundary element method in the external {\it solvent} region. As a proof-of-concept exercise, we used the simplest methods available: Johnson-N\'ed\'elec coupling with mass matrix and diagonal preconditioning, implemented in the Bempp-cl and FEniCSx libraries from a simple Python interface. We showcased our implementation by computing the polar component of the solvation free energy of a set of molecules using a constant and a Gaussian-varying permittivity. First, we validated our implementation against the finite differece code APBS (to 0.5\%), and then, we show scaling from protein G B1 (955 atoms) up to immunoglobuling G (20 148 atoms). For small problems we saw that the coupled method was efficient, outperforming a purely boundary integral approach, but scaled worse with size. For Gaussian-varying permittivities, which are beyond the applicability of boundary elements, we were able to run medium to large problems on a single workstation. However, our results illustrate the need to improve conditioning and incorporate distributed memory parallelism as we aim at larger systems. We hope this work will serve as inspiration for future developments that consider space-varying field parameters, and mixed linear-nonlinear schemes for molecular electrostatics with implicit solvent models. 
\end{abstract}

>>>>

\begin{abstract}
The Poisson-Boltzmann equation is widely used to model electrostatics in molecular systems. Available software packages solve it using finite difference, finite element, and boundary element methods, where the latter is attractive due to the accurate representation of the molecular surface and partial charges, and exact enforcement of the boundary conditions at infinity. However, the boundary element method is limited to linear equations and piece-wise constant variations of the material properties. In this work, we present a scheme that couples finite and boundary elements for the Poisson-Boltzmann equation, where the finite element method is applied in a confined {\it solute} region, and the boundary element method in the external {\it solvent} region. As a proof-of-concept exercise, we used the simplest methods available: Johnson-N\'ed\'elec coupling with mass matrix and diagonal preconditioning, implemented in the Bempp-cl and FEniCSx libraries from a simple Python interface. We showcased our implementation by computing the polar component of the solvation free energy of a set of molecules using a constant and a Gaussian-varying permittivity. First, we validated our implementation against the finite difference code APBS (to 0.5\%), and then, we show scaling from protein G B1 (955 atoms) up to immunoglobulin G (20 148 atoms). For small problems we saw that the coupled method was efficient, outperforming a purely boundary integral approach, but scaled worse with size. For Gaussian-varying permittivities, which are beyond the applicability of boundary elements, we were able to run medium to large problems on a single workstation. However, our results illustrate the need to improve conditioning and incorporate distributed memory parallelism as we aim at larger systems. We hope this work will serve as inspiration for future developments that consider space-varying field parameters, and mixed linear-nonlinear schemes for molecular electrostatics with implicit solvent models. 
\end{abstract}

### 2023-04-27 methods.tex
worked on results section
Additional comment while reading
Typo and references
resolved last comments and added acknowledgements
Updated Results with tol 1e-4

Regularized versions of Equation \eqref{eq:pbe} \cite{LuZhouHolstMcCammon2008,LeeGengZhao2021} are widely used to numerically solve the Poisson-Boltzmann equation with finite element or finite difference methods, and have recently been extended to super-Gaussian permitivitties in the solute.\cite{wang2022regularization} However, here we use the standard formulation in Equation \eqref{eq:pbe}, as it offers more flexibility when dealing with, for example, variable permitivitties, beyond super-Gaussian descriptions.

A common quantity of interest in implicit solvent models is the solvation free energy, which the change in Gibbs free energy as the molecule moves from vacuum into the solvent. Considering the charge distribution $\rho$ consists of point charges, this can be calculated as
%
\begin{equation}\label{eq:dG} 
\Delta G_{solv} = \tfrac{1}{2}\int_{\Omega_1} \rho(\mathbf{x})\phi_{r}(\mathbf{x}) = \tfrac{1}{2}\sum_{k=1}^{N_q} q_k\phi_r(\mathbf{x_k})
\end{equation}

@@

Several possible boundary integral formulations of Equation \eqref{eq:pbe} exist.~\cite{search2022towards} The simplest form was presented by Yoon ºand Lenhoff in 1991,~\cite{YoonLenhoff1990} known as the \emph{direct} formulation, however, the resulting system is usually ill-conditioned. Here, we use the better-conditioned formulation presented by Lu and co-workers,~\cite{LuETal2006,LuETal2009,debuhr2016dashmm} which reads
%
\begin{align}\label{eq:lu}
    & \tfrac{\phi_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) - \left(K_Y^\Gamma - \tfrac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right)\phi_2 + \left(V_Y^\Gamma - V_L^\Gamma\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}
     \nonumber \\
    &\tfrac{\epsilon_1}{\epsilon_2}\left(W_Y^\Gamma - W_L^\Gamma\right)\phi_2 +  \tfrac{\lambda_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) + \left(\tfrac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{x}}\left(\frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}\right)
\end{align}
%
In Equation \eqref{eq:lu}, $\phi_1 = \phi_1(\mathbf{x}_\Gamma)$ and $\phi_2 = \phi_2(\mathbf{x}_\Gamma)$ are the potentials on $\Gamma$ as we approach from $\Omega_1$ and $\Omega_2$, respectively. For convenience, we call the normal derivative on the interface ({\it ie.} the Neumann trace) $\lambda=\left(\frac{\partial}{\partial \mathbf{n}}  \phi_{\Gamma}  \right)$. $V$, $K$, $K^{\prime}$, and $W$ are the single-layer, double-layer, adjoint double-layer, and hypersingular operators for the Laplace (subscript $L$) and Yukawa (subscript $Y$, also known as modified Helmholtz) kernels. These are defined %in Equation~\eqref{eq:single_double} and 
as
%
\begin{align}\label{eq:all_op}
V_i^\Gamma \varphi (\mathbf{x}) &= \oint_\Gamma g_i(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
K_i^\Gamma \varphi (\mathbf{x}) &= \oint_\Gamma \frac{\partial g_i}{\partial\mathbf{n}'}(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}',\nonumber\\
K^{\prime\Gamma}_i\varphi (\mathbf{x}) &= \oint_\Gamma \frac{g_i}{\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
W^\Gamma_i\varphi (\mathbf{x}) &= - \oint_\Gamma \frac{\partial^2 g_i}{\partial\mathbf{n}'\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'
\end{align}
%
where $i \in \{L,Y\}$, $\varphi(\mathbf{x})$ can be any distribution over $\Gamma$, and %Green's functions $g_L, g_Y$ were defined in Equation~\eqref{eq:green_func}.
\begin{align}\label{eq:green_func}
g_L(\mathbf{x},\mathbf{x}')=\frac{1}{4\pi|\mathbf{x}-\mathbf{x}'|} \nonumber \\
g_Y(\mathbf{x},\mathbf{x}')=\frac{e^{-\kappa|\mathbf{x}-\mathbf{x}'|}}{4\pi|\mathbf{x}-\mathbf{x}'|}
\end{align}
are the corresponding free-space Green's functions.

@@

Having computed the electrostatic potential with Eq. \eqref{eq:lu}, we obtain the reaction potential ($\phi_r$) in Eq. \eqref{eq:dG} by subtracting out the Coulombic component out of $\phi_1$. This gives\cite{CooperBardhanBarba2014}
%
\begin{align} \label{eq:phi_reac}
\phi_{r}(\mathbf{r}) = - K_{L}^{\mathbf{r}} \phi_1 + V_{L}^{\mathbf{r}}  \lambda_1  ,
\end{align}
%
where $\mathbf{r}\in\Omega_1$.
This is valid as long as $\epsilon_1$ is a constant. We use Eq.~\eqref{eq:phi_reac} for both BEM-BEM and FEM-BEM coupling simulations with constant $\epsilon_1$.

@@

The boundary element method (BEM) is a standard tool for the numerical solution of the Poisson-Boltzmann equation in molecular electrostatics.~\cite{ZauharMorgan1985, Shaw1985} This was implemented in numerous codes, such as AFMPB,~\cite{LuETal2006} TABI,~\cite{GengKrasny2013} PyGBe,~\cite{CooperBardhanBarba2014,cooper2016pygbe} and more recently, with Bempp-cl.~\cite{search2022towards}. Its popularity comes because it reduces the dimension of the problem by using the boundary integral formulation. Unfortunately, this advantage comes at a cost: it requires a fundamental solution to be applied. There are many linear problems for which fundamental solutions are not known, for example, this is the case for the Poisson-Boltzmann equation with a heterogenous permittivity inside the molecule
%
   \begin{align} \label{eq:pbe_vp}
\nabla \cdot \left(\epsilon_1(\mathbf{x}) \nabla \phi_1(\mathbf{x})\right) &= \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k) \quad  \mathbf{x} \in \Omega_1\nonumber\\
\left(\nabla^2 - \kappa^2\right)\phi_2(\mathbf{x})  &= 0 \quad\mathbf{x}\in\Omega_2\nonumber\\
\phi_1(\mathbf{x})  = \phi_2(\mathbf{x}),  &\quad \epsilon_1(\mathbf{x})\frac{\partial\phi_1}{\partial\mathbf{n}}(\mathbf{x})  = \epsilon_2(\mathbf{x})\frac{\partial\phi_2}{\partial\mathbf{n}}(\mathbf{x})  \quad \mathbf{x}\in \Gamma. 
\end{align}
%
A finite element method (FEM) is more suitable for such case.

The coupling of finite element (FEM) and boundary element (BEM) methods is a widely used approach to solve multi-physical problems on unbounded domains [citation?], as it takes advantage of both methods. On one hand, the BEM satifies the infinite boundary conditions exactly when they decay to zero, and only approximates boundary conditions on surfaces, hence, it is commonly used for problems involving infinite or semi-infinite domains. On the other hand, the FEM is known for its robustness and universal applicability, even for problems of inhomogeneous or non-linear nature.
Here, we used the Johnson-N\'ed\'elec formulation,\cite{johnson1980coupling} which is the simplest form of FEM-BEM coupling, and we detail it next.

@@

This can also be written in matrix form. Let $\vec{\phi}_1 := [\phi_1^1, \dots, \phi_1^j]^T$ be the vector of canonical basis
functions of the finite element space $V_{h}^{j}$, %$\vec{\phi}_2 := [\phi_2^1, \dots, \phi_2^k]^T$ - the vector of canonical basis functions of $W_{h}^{k}$,  
$\vec{\lambda}_2 := [\lambda_2^1, \dots, \lambda_2^l]^T$ --- the vector of canonical basis
functions of $\Lambda_{h}^{l}$. % and $\vec{\widetilde{\phi}} := [\widetilde{\phi}^1, \dots, \widetilde{\phi}^m]^T$ be the vector of canonical basis functions of  $M_{h}^{m}$
% $T:\Omega_1 \rightarrow \Gamma$ and $E:\Gamma \rightarrow \Omega_1$ trace and extendable operators respectively.
We define the following matrices associated with the corresponding bilinear forms
\begin{align*}
A_{\alpha \beta} = \left(\nabla \phi_1^{\alpha}, \nabla \phi_1^{\beta} \right)_{\Omega_1}  
&,&
M_{\alpha \beta} = \left< \widetilde{\phi}^{\alpha}, \ \lambda^{\beta}_2 \right>_{\Gamma}, \\
K_{\alpha \beta} = \left<K_{Y}^{\Gamma} \gamma \phi^{\alpha}_1, \ \lambda^{\beta}_2 \right>_{\Gamma}
 &,&
V_{\alpha \beta} = \left<V_{Y}^{\Gamma} \lambda^{\alpha}_2, \ \lambda^{\beta}_2 \right>_{\Gamma}, 
\end{align*}
and vector associated with the corresponding linear form
\begin{equation*}
\vec{f}_{\beta} := \left(  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  \phi_1^{\beta} \right)_{\Omega_1}.
\end{equation*}
Using the above definition the discrete problem in Eq.~\eqref{eq:standard_fem_bem} can be written in the following matrix form
\begin{align*}
\begin{bmatrix}
\epsilon_1 A &  - \epsilon_1 M^T \\  
\left(\tfrac12 I + K \right) \gamma &  \tfrac{\epsilon_1}{\epsilon_2} V 
\end{bmatrix}
\begin{bmatrix}
\vec{\phi}_1 \\  
\vec{\lambda}_2
\end{bmatrix}
= 
\begin{bmatrix}
\vec{f} \\  
0
\end{bmatrix}.
\end{align*}

>>>>

Regularized versions of Equation \eqref{eq:pbe} \cite{LuZhouHolstMcCammon2008,LeeGengZhao2021} are widely used to numerically solve the Poisson-Boltzmann equation with finite element or finite difference methods, and have recently been extended to super-Gaussian permittivities in the solute.\cite{wang2022regularization} However, here we use the standard formulation in Equation \eqref{eq:pbe}, as it offers more flexibility when dealing with, for example, variable permittivities, beyond super-Gaussian descriptions.

A common quantity of interest in implicit solvent models is the solvation free energy, which is the change in Gibbs free energy as the molecule moves from the vacuum into the solvent. Considering the charge distribution $\rho$ consists of point charges, this can be calculated as
%
\begin{equation}\label{eq:dG} 
\Delta G_{solv} = \tfrac{1}{2}\int_{\Omega_1} \rho(\mathbf{x})\phi_{r}(\mathbf{x}) = \tfrac{1}{2}\sum_{k=1}^{N_q} q_k\phi_r(\mathbf{x_k})
\end{equation}

@@

Several possible boundary integral formulations of Equation \eqref{eq:pbe} exist.~\cite{search2022towards} The simplest form was presented by Yoon and Lenhoff in 1991,~\cite{YoonLenhoff1990} known as the \emph{direct} formulation, however, the resulting system is usually ill-conditioned. Here, we use the better-conditioned formulation presented by Lu and co-workers,~\cite{LuETal2006,LuETal2009,debuhr2016dashmm} which reads
%
\begin{align}\label{eq:lu}
    & \tfrac{\phi_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) - \left(K_Y^\Gamma - \tfrac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right)\phi_2 + \left(V_Y^\Gamma - V_L^\Gamma\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}
     \nonumber \\
    &\tfrac{\epsilon_1}{\epsilon_2}\left(W_Y^\Gamma - W_L^\Gamma\right)\phi_2 +  \tfrac{\lambda_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) + \left(\tfrac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{x}}\left(\frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}\right)
\end{align}
%\textcolor{blue}{Michal: We only have $\phi_2$ and $\lambda_2$ and the solution later is from $\phi_1$ and $\lambda_1$!!!}

%
In Equation \eqref{eq:lu}, $\phi_1 = \phi_1(\mathbf{x}_\Gamma)$ and $\phi_2 = \phi_2(\mathbf{x}_\Gamma)$ are the potentials on $\Gamma$ as we approach from $\Omega_1$ and $\Omega_2$, respectively. For convenience, we call the normal derivative on the interface ({\it ie.} the Neumann trace) $\lambda=\left(\frac{\partial}{\partial \mathbf{n}}  \phi_{\Gamma}  \right)$. $V$, $K$, $K^{\prime}$, and $W$ are the single-layer, double-layer, adjoint double-layer, and hypersingular operators for the Laplace (subscript $L$) and Yukawa (subscript $Y$, also known as modified Helmholtz) kernels. These are defined %in Equation~\eqref{eq:single_double} and 
as
%
\begin{align}\label{eq:all_op}
V_i^\Gamma \varphi (\mathbf{x}) &= \oint_\Gamma g_i(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
K_i^\Gamma \varphi (\mathbf{x}) &= \oint_\Gamma \frac{\partial g_i}{\partial\mathbf{n}'}(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}',\nonumber\\
K^{\prime\Gamma}_i\varphi (\mathbf{x}) &= \oint_\Gamma \frac{g_i}{\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
W^\Gamma_i\varphi (\mathbf{x}) &= - \oint_\Gamma \frac{\partial^2 g_i}{\partial\mathbf{n}'\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'
\end{align}
%
where $i \in \{L,Y\}$, $\varphi(\mathbf{x})$ can be any distribution over $\Gamma$, and %Green's functions $g_L, g_Y$ were defined in Equation~\eqref{eq:green_func}.
\begin{align}\label{eq:green_func}
g_L(\mathbf{x},\mathbf{x}')=\frac{1}{4\pi|\mathbf{x}-\mathbf{x}'|} \nonumber \\
g_Y(\mathbf{x},\mathbf{x}')=\frac{e^{-\kappa|\mathbf{x}-\mathbf{x}'|}}{4\pi|\mathbf{x}-\mathbf{x}'|}
\end{align}
are the corresponding free-space Green's functions.

@@

Having computed the electrostatic potential with Eq. \eqref{eq:lu}, we obtain the reaction potential ($\phi_r$) in Eq. \eqref{eq:dG} by subtracting out the Coulombic component out of $\phi_1$. This gives\cite{CooperBardhanBarba2014}
%
\begin{align} \label{eq:phi_reac}
\phi_{r}(\mathbf{r}) = - K_{L}^{\mathbf{r}} \phi_1 + V_{L}^{\mathbf{r}}  \lambda_1  ,
\end{align}
%
where $\mathbf{r}\in\Omega_1$, and we can obtain $\phi_1$ and $\lambda_1$ from $\phi_2$ and $\lambda_2$ by applying the interface conditions from Eq. \eqref{eq:pbe}.
This is valid as long as $\epsilon_1$ is a constant. We use Eq.~\eqref{eq:phi_reac} for both BEM-BEM and FEM-BEM coupling simulations with constant $\epsilon_1$.

@@

The boundary element method (BEM) is a standard tool for the numerical solution of the Poisson-Boltzmann equation in molecular electrostatics.~\cite{ZauharMorgan1985, Shaw1985} This was implemented in numerous codes, such as AFMPB,~\cite{LuETal2006} TABI,~\cite{GengKrasny2013} PyGBe,~\cite{CooperBardhanBarba2014,cooper2016pygbe} and more recently, with Bempp-cl.~\cite{search2022towards}. Its popularity comes because it reduces the dimension of the problem by using the boundary integral formulation. Unfortunately, this advantage comes at a cost: it requires a fundamental solution to be applied. There are many linear problems for which fundamental solutions are not known, for example, this is the case for the Poisson-Boltzmann equation with a heterogenous permittivity inside the molecule
%
   \begin{align} \label{eq:pbe_vp}
\nabla \cdot \left(\epsilon_1(\mathbf{x}) \nabla \phi_1(\mathbf{x})\right) &= \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k) \quad  \mathbf{x} \in \Omega_1\nonumber\\
\left(\nabla^2 - \kappa^2\right)\phi_2(\mathbf{x})  &= 0 \quad\mathbf{x}\in\Omega_2\nonumber\\
\phi_1(\mathbf{x})  = \phi_2(\mathbf{x}),  &\quad \epsilon_1(\mathbf{x})\frac{\partial\phi_1}{\partial\mathbf{n}}(\mathbf{x})  = \epsilon_2(\mathbf{x})\frac{\partial\phi_2}{\partial\mathbf{n}}(\mathbf{x})  \quad \mathbf{x}\in \Gamma. 
\end{align}
%
A finite element method (FEM) is more suitable for such a case.

The coupling of finite element (FEM) and boundary element (BEM) methods is a widely used approach to solve multi-physical problems on unbounded domains [citation?], as it takes advantage of both methods. On one hand, the BEM satisfies the infinite boundary conditions exactly when they decay to zero, and only approximates boundary conditions on surfaces, hence, it is commonly used for problems involving infinite or semi-infinite domains. On the other hand, the FEM is known for its robustness and universal applicability, even for problems of inhomogeneous or non-linear nature.
Here, we used the Johnson-N\'ed\'elec formulation,\cite{johnson1980coupling} which is the simplest form of FEM-BEM coupling, and we detail it next.

@@

This can also be written in matrix form. Let $\vec{\phi}_1 := [\phi_1^1, \dots, \phi_1^j]^T$ be the vector of canonical basis
functions of the finite element space $V_{h}^{j}$, %$\vec{\phi}_2 := [\phi_2^1, \dots, \phi_2^k]^T$ - the vector of canonical basis functions of $W_{h}^{k}$,  
$\vec{\lambda}_2 := [\lambda_2^1, \dots, \lambda_2^l]^T$ --- the vector of canonical basis
functions of $\Lambda_{h}^{l}$. % and $\vec{\widetilde{\phi}} := [\widetilde{\phi}^1, \dots, \widetilde{\phi}^m]^T$ be the vector of canonical basis functions of  $M_{h}^{m}$
% $T:\Omega_1 \rightarrow \Gamma$ and $E:\Gamma \rightarrow \Omega_1$ trace and extendable operators respectively.
We define the following matrices associated with the corresponding bilinear forms
\begin{align*}
A_{\alpha \beta} = \left(\nabla \phi_1^{\alpha}, \nabla \phi_1^{\beta} \right)_{\Omega_1}  
&,&
M_{\alpha \beta} = \left< \widetilde{\phi}^{\alpha}, \ \lambda^{\beta}_2 \right>_{\Gamma}, \\
K_{\alpha \beta} = \left<K_{Y}^{\Gamma} \gamma \phi^{\alpha}_1, \ \lambda^{\beta}_2 \right>_{\Gamma}
 &,&
V_{\alpha \beta} = \left<V_{Y}^{\Gamma} \lambda^{\alpha}_2, \ \lambda^{\beta}_2 \right>_{\Gamma}, 
\end{align*}
and vector associated with the corresponding linear form
\begin{equation*}
\vec{f}_{\beta} := \left(  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  \phi_1^{\beta} \right)_{\Omega_1}.
\end{equation*}
Using the above definition the discrete problem in Eq.~\eqref{eq:standard_fem_bem} can be written in the following matrix form
\begin{align}\label{eq:fembem_matrix}
\begin{bmatrix}
\epsilon_1 A &  - \epsilon_1 M^T \\  
\left(\tfrac12 I - K \right) \gamma &  \tfrac{\epsilon_1}{\epsilon_2} V 
\end{bmatrix}
\begin{bmatrix}
\vec{\phi}_1 \\  
\vec{\lambda}_2
\end{bmatrix}
= 
\begin{bmatrix}
\vec{f} \\  
0
\end{bmatrix}.
\end{align}
